---
title: "R Notebook"
output: 
html_documnet: default
git_document: default
---

#Reading in the Data

```{r}
library(tidyverse)
library(dplyr)
library(lazyeval)

#read in Washington Post data and convert into df with fatalities by state, with 
#ethnicity as rows

fatal_shootings <- read_csv('fatal_shootings.csv', col_names = TRUE)

state_shootings <- count(fatal_shootings, race, state)

shootings <- spread(state_shootings, key = state, value = n)

##This dataframe is FULL of NA's the better to test our sum function
```


#Functions and error handling

Write a function that fulfills the following criteria:

1.  tidyverse compatible (i.e., the first argument must be a data frame)

2.  adds two arbitrary columns of and put them in a new column of that data frame, with     the name of the new column specified by the user

3.  warning if any invalid arguments are provided. Invalid arguments might include:
        The first argument is not a data frame
        Less than two valid columns are specified to add (e.g., one or both of the                  column names isn't in the supplied data frame)
        The columns specified are not numeric, and therefore can't be added - use       
            tryCatch() for this
    If the columns to add aren't valid but the new column name is, the function should      create a column of NA values

```{r}
library(dplyr)
library(lazyeval)

df.sum <- function(df.in, col.1, col.2, name)  {

  if(is.data.frame(df.in) != TRUE){               #warning if first arg not df
    warning('df.in is not a dataframe')}

  if(is.numeric(lazyeval::lazy_eval(substitute(col.1), df.in)) != TRUE){                
    warning('col.1 is not a numeric vector')}     

  if(is.numeric(lazyeval::lazy_eval(substitute(col.2), df.in)) != TRUE){
    warning('col.2 is not a numeric vector')}     #warning if col not numeric 

  dots <- setNames(list(lazyeval::interp(~sum(x, y, na.rm = TRUE),
                                         x = substitute(col.1), y = substitute(col.2))),
                   name)

  df.out <- rowwise(df.in) %>%
    mutate_(.dots = dots)

  df.out 
}


df.sum(shootings, RI, CT, 'northeast')

```


#Loop and performance metric tasks

Write a function that uses a for loop to calculate the sum of the elements of a vector, which is passed as an argument (i.e., it should do the same thing that sum() does with vectors). 

your_fun(1:10^4) should return 50005000.

```{r}
a <- (1:10^4)

vec_sum <- function(a){

n <- as.numeric(length(a))

vec.sum = 0

for (i in 1:n) {
  vec.sum <- vec.sum + a[i]
               }

print(vec.sum)

}
```


Use the microbenchmark::microbenchmark function to compare the performace of your function to that of sum in adding up the elements of the vector 1:10^4.
    
```{r}
library(microbenchmark)

test.vec <- 1:10^4
microbenchmark(
    vec_sum(test.vec),
    sum(test.vec)
    )
```

Is there a difference? Why?

Obviously, the function I wrote using a for-loop takes longer to execture than the sum function. Without knowing too much about what is going on when we deploy R's sum() function
under the hood, as it were, the user-defined function vec_sum() uses multiple functions:
as.numeric()
length()
and then, of, course the for-loop. 

Presumably, R's sum() function is going to be more efficent than this. 

