---
title: 'Task 7'
output: 
html_documnet: default
git_document: default
---

Completed draft: 3 March 17
First upload: 4 March 17

```{r}
library(tidyverse)
library(dplyr)
library(nycflights13)
```
#Section I

Using nycflights13::weather, determine whether there are any clear outliers in wind speed (wind_speed) that should be rejected. If so, filter those bad point(s) and proceed.

```{r}
#plot to visiualize outlier
nycflights13::weather %>% ggplot(aes(x = origin, y = wind_speed)) + geom_boxplot()

#filter out values we don't want
nycflights <- nycflights13::weather %>% filter(wind_speed < 250.0) 

#plotting again to see if additional outliers
nycflights %>% ggplot(aes(x = origin, y = wind_speed)) + geom_boxplot()
```

Make a table and a plot of median wind speed by direction, for each airport. Optional fun challenge: If you like, this is a rare opportunity to make use of coord_polar().

```{r}
#create new tibble with origin, wind direction, and median wind speed by direction
nycwind <- nycflights %>% 
                    group_by(origin, wind_dir) %>%
                    filter(wind_speed > 0) %>%
                    summarise(med_sp = median(wind_speed)) %>%
                    mutate(med_sp) %>%
                    select(origin, wind_dir, med_sp)

#point plot
nycwind %>% ggplot(aes(x = wind_dir, y = med_sp, color = origin)) +
      geom_point()

#plot using polar coordinates
nycwind %>% ggplot(aes(x = med_sp, y = wind_dir, color = origin)) +   
      geom_point() +
      coord_polar(theta = 'y', start = -pi/2, direction = -1) 
```


What direction has the highest median speed at each airport?

```{r}
#this tibble displays the direction(s) of the max medium wind speeds by airport
max_med_sp <- nycwind %>%
              group_by(origin) %>%
              filter(med_sp == max(med_sp))

max_med_sp
```

#Section II

Using nycflights13::flights and nycflights13::airlines, make a table with two columns: airline name (not carrier code) and median distance flown from JFK airport. The table should be arranged in order of decreasing mean flight distance. Hint: use a _join function to join flights and airlines.

```{r}
#visualize the dfs we're starting with
nycflights13::flights
nycflights13::airlines

#combine the dfs matching carrier with name
airline_dist_jfk <- (nycflights13::flights [,c('carrier', 'distance')] %>%
                    left_join(nycflights13::airlines, by = 'carrier'))[,-1]

#get med, mean and put in descending order
airline_dist_jfk <- airline_dist_jfk %>%
                    group_by(name) %>%
                    summarise(med_dist = median(distance), mean_dist = mean(distance)) %>%
                    arrange(desc(mean_dist))

airline_dist_jfk
```

Make a wide-format data frame that displays the number of flights that leave Newark ("EWR") airport each month, from each airline


```{r}
Newark_flights <- nycflights13::flights %>%
                  left_join(nycflights13::airlines, by = 'carrier') %>%    #get carrier name
                  filter(origin == 'EWR') %>%
                  group_by(month, name) %>%
                  summarise(num_flight = length(flight)) %>%    #col length of flight is number of flights
                  spread (month, num_flight)                               #orient horizontally

Newark_flights
```

#Section III

Using the babynames dataset, identify the ten most common male and female names in 2014. Make a plot of their frequency (prop) since 1880.

```{r}
babynames::babynames
#already in descending order of popularity by year

#new tibble with names from 2014, with top 10 rows for each sex
top_names_14 <- babynames::babynames %>%
                filter(year == 2014) %>%
                group_by(sex) %>%
                top_n(10,n)

#new tiddle with values of names top 10 names for M and F and filter for only those rows in babynames
top_names_freq <- babynames::babynames %>%
                filter(paste0(sex, name) %in% paste0(top_names_14$sex, top_names_14$name))

#plot year,prop, color to distinguish name
top_names_freq %>% ggplot(aes(x = year, y = prop, color = name)) +
                geom_line() +
                facet_grid(.~sex)

```

Make a single table of the 26th through 29th most common girls names in the year 1896, 1942, and 2016

```{r}
#there's a short way to do this, but this is mine so far

#tibble of 1896 F names in descending order of popularity
female_1896 <- babynames::babynames %>%
                filter(sex == 'F') %>%
                filter(year == 1896)
#select rows 26-29 correspond to 26th-29th most popular names
female_1896 <- female_1896[26:29,]          #table of 26-29th popular names 1896

female_1942 <- babynames::babynames %>%
                filter(sex == 'F') %>%
                filter(year == 1942)

female_1942 <- female_1942[26:29,]          #table of 26-29th popular names 1942

female_2014 <- babynames::babynames %>%
                filter(sex == 'F') %>%
                filter(year == 2014)

female_2014 <- female_2014[26:29,]          #table of 26-29th popular names 2014

#join the 3 tibbles two at a time
female_26_29 <- full_join(female_1896, female_1942)
                
female_26_29 <- full_join(female_26_29, female_2014)  

female_26_29
```

Section IV


Write task that involves some of the functions on the Data Wrangling Cheat Sheet and execute it. You may either use your own data or data packages (e.g., the ones listed here).

Use the timeseries sunspot.year to visualize and locate peak sunspot activity years.

```{r}
library(datasets)
library(pastecs)
datasets::sunspot.year

#build tibble from the ts sunspot.years
sunspot <- data_frame(year = 1700:1988, spots = as_vector(sunspot.year))
tbl_df(sunspot)

#plot time, activity
sunspot %>% ggplot(aes(x = year, y = spots)) +
                    geom_line()

#use pastecs library function turnpoints to find local maxima, minima
e <- turnpoints(sunspot.year)

#create dataframe of peak and year
peaks <- data_frame(peak = as_vector(e$peaks), s = as_vector(e$points))

#the turnpoints function omitted a point when there were 2 consecutive 0's at 
#1711 and 1712. Easier to omit a row in sunspots than add one in peaks
sunspot <- sunspot %>% filter(year != 1711) 

#combine data into dateframe
cycles <- bind_cols(sunspot, peaks)

#convert to tibble
tbl_df(cycles)

maxima <- cycles %>% filter(peak == TRUE) %>%
                    select(year, spots)

#cyclic maximum sunpot activity years
maxima
```

